---
title: "reports"
author: "Matthias Spichiger"
date: "2023-10-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## epoxy in Reports

### Setup

```{r}
library(epoxy)
```


Loading epoxy adds four new [knitr engines](https://bookdown.org/yihui/rmarkdown/language-engines.html), or chunk types.
Each type lets you intermix text with R code or data (`expr` in the table below), and each is geared toward a different output context.

| Engine      | Output Context         | Delimiter                                                 |
| epoxy       | all-purpose markdown   | `{expr}`                                                  |
| epoxy_html  | HTML                   | `{{expr}}`                                                |
| epoxy_latex | LaTeX                  | `<<expr>>`                                                |
| whisker     | all-purpose            | [mustache template language](https://mustache.github.io/) |

**Caution:** Previously, epoxy provided a `glue`engine, but this conflicts with a similar chunk engine by the `glue`package.
You can update existing documents to use the `epoxy`engine, or you can explicitly use epoxy's `glue`chunk by including the following in your setup chunk:

`use_epoxy_glue_engine()`


### Using epoxy chunks 

To use epoxy in your R Markdown document, create a new chunk using the engine of your choice.
In that chunk, write in markdown, HTML, or LaTeX as needed, wrapping R expressions inside the delimiters for the epoxy chunk.

```{epoxy}
The average speed of the cars was **{mean(cars$speed)} mph**.
But on average the distance traveled was only _{mean(cars$dist)}_.
```

epoxy is built around `glue::glue()`, which evaluates the R expression inside the `{}` and inserts the results into the string.
The chunk above is equivalent to this call to `glue::glue()`:

```{glue}
glue::glue("The average speed of the cars was **{mean(cars$speed)} mph**.
But on average the distance traveled was only _{mean(cars$dist)} ft_.")
```


### Inline transformers

**epoxy** provides inline transformers inspired by [cli's inline markup](https://cli.r-lib.org/reference/inline-markup.html).
This transformer is enabled by default in `epoxy()`, `epoxy_html()` and `epoxy_latex()` and their respective knitr chunk engines.

Here is an example using a small list containing data about a movie (expand the section below to see the full code). We can use the inline transformer to format the replacement text as we build up a description from this data.

```{r}
movie <- list(
  year = 1989,
  title = "Back to the Future Part II",
  budget = 4e+07,
  domgross = 118450002,
  imbd_rating = 7.8,
  actors = c(
    "Michael J. Fox",
    "Christopher Lloyd",
    "Lea Thompson",
    "Thomas F. Wilson"
  ),
  runtime = 108L
)
```

```{epoxy}
The movie {.emph {.titlecase movie$title}}
was released in {.strong movie$year}.
It earned {.dollar movie$domgross}
with a budget of {.dollar movie$budget},
and it features movie stars
{.and movie$actors}
```

Read more about inline transformations in `?epoxy_transform_inline`.


### Transform replaced values

You can use the `epoxy_transform_wrap()`with the `epoxy_transform` chunk option to wrap the evaluated R expression in formatting or templating text.

Or you can use the pre-set `epoxy_transform_bold()`, `epoxy_transform_italic()`, or `epoxy_transform_code()` transformers or with `epoxy_transform()`.

```{epoxy, .transformer = epoxy_transform("bold")}
All cars stopped between {min(cars$dist)} and {max(cars$dist)} feet
from a starting speed of {min(cars$speed)}---{max(cars$speed)}
```


### epoxy chunks are vectorized

Unlike inline R code, the epoxy chunks are vectorized. This can be something to watch out for or it can be an advantage:

```{epoxy}
{1:4}. "{letters[1:4]}" is for {c("apple", "banana", "coconut", "donut")}
```

You can collapse fields automatically using the `epoxy_transform_collapse()`transformer.

You can then choose how vectors are collapsed by adding `*`, `&` or `|` to the end of the expression.

- `*` collapses with commas, e.g. `{letters[1:3]*}`.
- `&`collapses with commas and adds `" and "` between the last two items.
- `|` collapses with commas and adds `" or "` between the last two items.

```{epoxy, .transformer = epoxy_transform("collapse")}
- The first three letters are {letters[1:3]*}.
- When capitalized, they are {LETTERS[1:3]&}.
- They are indexed by {1:3|}
```

You can change the separator between entries and between the last entry using the `sep`, `last` and the `_and` and `_or` specific arguments of the `epoxy_transform_collapse()` function.


### Templating with epoxy chunks 

It is possible to create a reusable template. Use the `ref.label` chunk option to reusa a template using the values in the `.data` chunk option, which can be a list or a data frame.
```{r}
mpg <- data.frame(
  manufacturer = c("Chevrolet", "Dodge", "Ford"),
  model = c("Malibu", "Caravan", "Expedition"),
  cty = c(19, 7, 11),
  hwy = c(27, 24, 17)
)
```

```{epoxy, car-name, eval=FALSE}
- A {manufacturer} {model} gets {cty} city and {hwy} highway miles per gallon.
```

```{epoxy, ref.label="car-name", .data = mpg}

```


### Whisker engine

Sometimes the `epoxy` engine does not quite deliver the template power you need. In these cases, you can use the whisker engine instead.

```{r}
contestant <- list(name = "R User", value = 1000, taxed = 600, in_ca = TRUE)
```

```{whisker .data = contestant, echo=FALSE}
Hello {{name}}:
You have just won ${{value}}!
{{#in_ca}}
Well, ${{taxed}}, after taxes.
{{/in_ca}}
```

```{r}
contestant <- list(name = "R User", value = 1000, taxed = 600, in_ca = TRUE)
```


### HTML and LaTeX chunks

#### Markdown chunks

The `epoxy` engine can be used in any output format. In practice, it works best in markdown (i.e. generally in R Markdown or Quarto)

```{epoxy, .data = mpg}
- **{manufacturer}** _{model}_
```

If you are writing for an HTML or LaTeX output, however, you may need to write literal HTML or LaTeX in your document. With the `epoxy`chunk, you would need to escape any `{` or `}` in your text by doubling them, otherwise the content will be treated as a template expression. To avoid this friction, epoxy provides two additional chunk engines, `epoxy_html`for writing raw HTML and `epoxy_latex`for writing raw LaTeX.


#### Raw HTML chunks

Use the `epoxy_html` block to epoxy (glue) R and HTML together. The output is [raw HTML](https://pandoc.org/MANUAL.html#raw-htmltex). By default, expressions in these types of blocks are wrapped in `{{` and `}}`, like the whisker templates above.

<ul>
```{epoxy_html, .data = mpg}
  <li><string>{{manufacturer}}</strong> <em>{{model}}</em></li>
```
</ul>


Notice that the output is HTML but wrapped in a [pandoc raw HTML block](https://pandoc.org/MANUAL.html#extension-raw_attribute), which tells pandoc that the content is HTML that should not be modified. It also means that the output of the chunk will [only be included in HTML documents](https://bookdown.org/yihui/rmarkdown-cookbook/raw-content.html).

If your `epoxy_html`block is contained within another raw html block, or if you want to force the output to appear, you can set the chunk `hmtl_raw = FALSE`

```{=html}
<ul>
```{epoxy_html, .data = mpg, html_raw = FALSE}
  <li><string>{{manufacturer}}</strong> <em>{{model}}</em></li>
```

</ul>
```
`epoxy_html` uses two custom transformers, `epoxy_transform_inline()`and `epoxy_transform_html()`, applying the html transformer before the inline transformer. With `epoxy_transform_html()` you can use `element.class#id`syntax to wrap expressions in HTML elements (all parts are optional).

Let's use this syntax to place manufacturer and model in `<strong>`and `<em>` elements, each with a custom class.

```{=latex}
<ul>
```{epoxy_html, .data = mpg}
<li>
  {{strong.car-make manufacturer}}
  {{em.car-model model}}
</li>
```
</ul>

<ul>
```{=html}
  <li>
    <strong class="car-make">Chevrolet</strong><strong class="car-make">Dodge</strong><strong class="car-make">Ford</strong>
    <em class="car-model">Malibu</em><em class="car-model">Caravan</em><em class="car-model">Expedition</em>
  </li>
```
</ul>
```
Because the `epoxy_transform_html()`transformer uses `.<class>` to create `<span class="class">`elements, `epoxy_html()`also recognizes `@<inline>`to access the inline transformers. So `{{.uppercase manufacturer}}` is assumed to be a CSS class and not an inline transformer class.

```{epoxy_html, .data = mpg[1, ]}
{{.uppercase manufacturer}}

{{@uppercase manufacturer}}
```

```{=html}
<span class="uppercase">Chevrolet</span>

CHEVROLET
```


### Raw LaTeX chunks

Similarly, you can also use `epoxy_latex`chunks to epoxy R and LaTeX together. Wrap expressions in these types of chunks with `<<` and `>>`.

```{=latex}
\begin{itemize}
```{epoxy_latex, .data = mpg}
\item <<.strong manufacturer>> <<.emph model>> gets <<cty>> city and <<hwy>> highway miles per gallon.
```
\end{itemize}
```

In R Markdown knitting to LaTeX, this renders as:

```{=latex}
\begin{itemize}
```{=latex}
\item \textbf{Chevrolet} \emph{Malibu} gets 19 city and 27 highway miles per gallon.
\item \textbf{Dodge} \emph{Caravan} gets 7 city and 24 highway miles per gallon.
\item \textbf{Ford} \emph{Expedition} gets 11 city and 17 highway miles per gallon.
```
\end{itemize}
```

Note that, like `epoxy_html_chunks`, `epoxy_latex` places the output in [raw latex](https://bookdown.org/yihui/rmarkdown-cookbook/raw-latex.html) blocks. This behavior can be disabled by setting the chunk option `latex_raw = FALSE`.





