---
title: "shiny"
author: "Matthias Spichiger"
date: "2023-10-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Epoxy in Shiny

```{r}
library(epoxy)
```

### Templating in Shiny

[Shiny apps](https://shiny.posit.co/) are a great way to design interactive web applications and epoxy includes several functions to help you weave reactive data into your apps.

Here are some ways you can use epoxy in your Shiny apps:

  1. Make the text portion of any element in your Shiny UI update dynamically.

  2. Weave reactive text into prose in your app.

  3. Build powerful templates using the [mustache templating language](https://mustache.github.io/).
  
Shiny already includes two reactive text outputs:

- `shiny::uiOutput()` (a.k.a. `shiny::htmlOutput()`) and
- `shiny::textOutput()`.

These are great for displaying reactive text in your app, but they have some limitations:

- `uiOutput()`tends to move your UI code into the `server`function, making it harder to know the final structure of your UI.

- `textOutput()`is great for displaying reactive text, but it takes some work to get the spacing around the dynamic text *just right*.

epoxy improves the experience of writing apps with dynamic text and templates.


## Introdcuing epoxy in Shiny

### A basic shiny app with textOutput()

Let's start with an example Shiny app. It's a simple but friendly app that greets the user by name.

```{r}
library(shiny)

ui <- fluidPage(
  textInput("name", "What's your name?"),
  p("Hello,", textOutput("greeting", inline = TRUE), "!")
)

server <- function(input, output, session) {
 output$greeting <- renderText(input$name) 
}

shinyApp(ui, server)
```

This gets you pretty close to what you want, but you have to remember to include `inline = TRUE` in `textOutput()`.
There's also some extra space between the user's name and the exclamation point that you'd probably like to get rid of.

## Setting up `ui_epoxy_html()`

Here's how to approach dynamic text with `ui_epoxy_html()`:

  1. Wrap a portion of your UI in `ui_epoxy_html()` and give it an `.id`.
  
  2. Use `{{ name }}` syntax to define fields where the dynamic text should go.
  
  3. In your server code, assign `render_epoxy()` to an output matching the UI's `.id` and pass in the reactive data as arguments with names matching the dynamic fields.
  
```{r}
library(shiny)
library(epoxy)

ui <- fluidPage(
  textInput("name", "What's your name?"),
  ui_epoxy_html(          #<< Template wrapper
    .id = "greeting",     #<< Unique ID
    p("Hello, {{name}}!") #<< Dynamic text
  )                     
)

server <- function(input, output) {
  output$greeting <- render_epoxy( #<< Connect to template
    name = input$name              #<< Reactive data field
  )
}

shinyApp(ui, server)
```


## Default or error values

Another advantage of using `us_epoxy_html()` over `textOutput()`is that you can set default values that appear immediately while your app is loading or that are used when an error occurs.

In the next app, `name`is set to `"friend"` by default in `ui_epoxy_html()`, and on the side I've also added a `validate()`call indicating that we need a name with at least 2 characters.

```{r}
library(shiny)
library(epoxy)

ui <- fluidPage(
  textInput("name", "What's your name?"),
  ui_epoxy_html(
    .id = "greeting",
    p("Hello, {{name}}!"),
    name = "friend"
  )
)

server <- function(input, output) {
  name <- reactive({
    validate(need(
      nchar(input$name) > 2,
      "Name must be more than 2 characters."
    ))
    input$name
  })
  
  output$greeting <- render_epoxy(
    name = name()
  )
}

shinyApp(ui, server)
```

If the user has not yet entered a name of more than 2 characters, the text for the `name`field will use the default value and will have a red squiggle below it. Hovering over the squiggle reveals the error message.


## A few more things about `us_epoxy_html()`

First, you can reference the same reactive value, e.g. `{{ name }}`, as many times as you want in your template. This value can also go just about anywhere in your UI as long as it's okay to put a `<span>`around the text.

```{r}
ui <- fluidPage(
  textInput("name", "What's your name?"),
  ui_epoxy_html(
    .id = "greeting",
    selectInput(
      inputId = "color",
      label = "What's your favorite color, {{ name }}?",
      choices = c("red", "green", "blue", "purple", "yellow")
    ),
    name = "friend"
  )
)

server <- function(input, output) {
  output$greeting <- render_epoxy(
    name = input$name
  )
}

shinyApp(ui, server)
```


You can use the `{{ <markup> <expr> }}` syntax from `epoxy_html()`, which makes it possible to determine which HTML element and class is used to contain the dynamic text. If you send an array of values to this reactive field, the tag is used as a template, making it easy to do things like dynamically update a list.

```{r}
library(shiny)
library(epoxy)

ui <- fluidPage(
  textInput("faves", "What are your favorite fruits?"),
  helpText("Enter a liist of comma-separated fruits."),
  ui_epoxy_html(
    .id = "fruit_list",
    tags$ul("{{ li fruits }}"),
    fruits = "favorite fruits"
  )
)

server <- function(input, output) {
  fruits <- reactive({
    validate(need(
      nzchar(input$faves),
      "Please share your favorite fruits."
    ))
    fruits <- trimws(strsplit(input$faves, ",\\s*")[[1]])
    fruits[nzchar(fruits)]
  })
  
  output$fruit_list <- render_epoxy(
    fruits = fruits()
  )
}

shinyApp(ui, server)
```


Three more quick things about `ui_epoxy_html()`:

  1. It assumes that bare character strings are HTML, so you don't have to worry about adding `HTML()` all over the place.
  
  2. The replacement text is *not assumed to be HTML*, by default, to save you from accidentally injecting unsafe HTML from user input into your app. If you're very certain that a field will only contain safe HTML, you can mark it as safe for HTML with three braces, e.g. `{{{ <expr> }}}`.
  
  3. The replacement text is sent as bare text or HTML, making it more like `textOutput()` than `uiOutput()`. In particular, where `uiOutput()`would allow you to send arbitrary widgets based on [htmlwidgets](https://github.com/ramnathv/htmlwidgets) or [htmltools](https://github.com/rstudio/htmltools), `ui_epoxy_html()` only ever sends the bare text or HTML.
  
  
## Connecting epoxy with a reactive data frame

One of my favorite use cases for epoxy's Shiny functions is to crate a UI template that's filled in by row in a data frame. IN this pattern, the app's inputs are combined in a reactive expression that filters the data frame down to a single row. Then, that row is sent via `render_epoxy()` to the UI, where it's dynamically injected into the template UI.

Here's a small example using epoxy's built in `bechdel` data set, a small data set with the 10 highest-rated movies that pass the [Bechdel test](https://en.wikipedia.org/wiki/Bechdel_test). In this app, the user picks a movie and the template below is filled out with information from the data set for that movie.

```{r}
library(shiny)
library(epoxy)

movie_choices <- bechdel$imdb_id
names(movie_choices) <- bechdel$title

ui <- fixedPage(
  selectInput("movie", "Pick a movie", choices = movie_choices),
  ui_epoxy_html(
    .id = "movie_info",
    p(
      "{{ em title }} was released",
      "in {{ strong year }}.",
      "It was directed by {{ director }}",
      "and was rated {{ rated }}."
    )
  )
)

server <- function(input, output, session) {
  movie <- reactive({
    # Use the inputs to filter a single row
    bechdel[bechdel$imdb_id == input$movie, ]
  })
  
  # Pass the reactive data frame to 
  # the .list argument of render_epoxy()
  output$movie_info <- render_epoxy(.list = movie())
}

shinyApp(ui, server)
```

Notice that instead of passing named arguments for each field to `render_epoxy()`, we pass the entire data frame to the `.list` argument.

```{r}
render_epoxy(.list = movie())
```

You can use this same pattern with a list in a `reactiveVal()`, a `recactive()` that returns a data frame, a list, or a list-like object, or a `reactiveValues()` list. And `.list`can coexist with named expressions.

```{r}
render_epoxy(
  name = input$name,
  .list = movie()
)
```

If you want to build the entire list within a single reactive expression, set `.list`equal to the expression, wrapped in braces:

```{r}
render_epoxy(.list = {
  list(
    name = input$name,
    age = input$age
  )
})
```


## Markdown templates

If you are using epoxy to write data-driven prose, you might want to use markdown for your templates, rather than writing in HTML.

`ui_epoxy_markdown()` is a version of `ui_epoxy_html()`that uses markdown syntax instead of HTML syntax.

Let's revisit our movie app from the example, but this time using markdown for the template.

```{r}
library(shiny)
library(epoxy)

movie_choices <- bechdel$imdb_id
names(movie_choices) <- bechdel$title

ui <- fixedPage(
  selectInput("movie", "Pick a movie", choices = movie_choices),
  ui_epoxy_markdown(
    .id = "movie_info",
    "_{{ title }}_ was released",
    "in **{{ year }}**.",
    "It was directed by {{ director }}",
    "and was rated {{ rated }}."
  )
)

server <- function(input, output, session) {
  movie <- reactive({
    bechdel[bechdel$imdb_id == input$movie, ]
  })
  
  output$movie_info <- render_epoxy(.list = movie())
}

shinyApp(ui, server)
```


For an even more involved example, try the epoxy markdown example app

```{r}
run_epoxy_example_app("ui_epoxy_markdown")
```


## Mustache templates

For more complex templates, you might want to use a template language like [Mustache](https://mustache.github.io/). In R, we know this syntax from the [whisker](https://github.com/edwindj/whisker) package.

`ui_epoxy_mustache()`wraps the Mustache language, letting you blend typical [shiny](https://shiny.posit.co/) and [htmltools](https://github.com/rstudio/htmltools) UI with the mustache template.

When would you use `ui_epoxy_mustache()` instead of `ui_epoxy_html()`?

- If your template variables are used as HTML attributes, e.g. in links or images (via the `href` or `src` attributes).

- If you want to use mustache's conditional logic, e.g. `{{#<expr>}} ... {{/<expr>}}.


Let's revisit our favorite fruits example app from earlier.

```{r}
library(shiny)
library(epoxy)

ui <- fluidPage(
  textInput("faves", "What are your favorite fruits?"),
  ui_epoxy_mustache(
    id = "fruit_list",
    tags$ul(
      # If fruits is not empty, wrap into list items
      "{{#fruits}}",
      tags$li("{{.}}"),
      "{{/fruits}}",
      # If fruits is empty, show a help message
      "{{^fruits}}",
      tags$li(
        class = "text-muted",
        "Enter a list of comma-separated fruits."
      ),
      "{{/fruits}}"
    )
  )
)

server <- function(input, output) {
  fruits <- reactive({
    req(input$faves)
    fruits <- trimws(strsplit(input$faves, ",\\s*")[[1]])
    fruits[nzchar(fruits)]
  })
  
  output$fruit_list <- render_epoxy(
    fruits = fruits()
  )
}

shinyApp(ui, server)
```

You can find a more detailed example in the epoxy mustache example app.

```{r}
run_epoxy_example_app("ui_epoxy_mustache")
```

One important thing to note about `ui_epoxy_mustache()`is that, unlike `ui_epoxy_html()`, the entire template is re-rendered (in the browser) whenever a reactive source updates. So it would be better to use smaller, localized templates that to wrap your entire app in `ui_epoxy_mustache()`.
